---
title: 'Computation: Three Birthday Problem'
author: "Austin Pesina"
date: "6/4/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```

## What is the probability that three or more people have a common birthday?


Let $k$ be the number of people. Exclude February 29 and assume the other 365 days are equally likely. Assume independence of births.

Let $k=365$.

$P(no match) = \frac{365*364*363*...*(365-k+1)}{365^k}$

$P(match) = {k\choose3}$

The probability of multiple people sharing a birthday will depend on the number of people in the room. R has a built in function under the `stats` library called `pbirthday` that we will use to compute the probability of coincidences.

Note: This function uses $n$ instead of $k$. It also allows us to change the number of coincidences.

```{r}
x <- seq(0, 300, by = 25)
for (i in x) 
  
print(paste0(pbirthday(i, coincident = 3)), sep = "")

pbirthday(88, coincident = 3)
```

By using a for loop, we see that when $k = 75$, our probability of 3 people sharing a birthday is $36.4\%$. When we increase to $k=100$, our probability jumps up to $63.9\%$. When $k=88$, we get our first value over $50\%$.

```{r}
a <- pbirthday(0, coincident = 3)
b <- pbirthday(50, coincident = 3)
c <- pbirthday(100, coincident = 3)
d <- pbirthday(150, coincident = 3)
e <- pbirthday(200, coincident = 3)
f <- pbirthday(250, coincident = 3)
g <- pbirthday(300, coincident = 3)

tab <- matrix(c(0, 50, 100, 150, 200, 250, 300, a, b, c, d, e, f, g), ncol = 2, byrow = F)
tab

```


```{r}
birthdays <- seq(0, 300, by = 50)
prob <- c(a,b,c,d,e,f,g)
df <- data.frame(birthdays, prob)

ggplot(data = df) +
         geom_line(mapping = aes(x = birthdays, y = prob))
```

As we can see, once we have $k=200$, we have an almost 100% probability of having 3 people with the same birthday.